---
title: "Test with genus abundance data "
author: "Weijia Xiong"
date: "5/14/2020"
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)
library(readxl)
library(tidyverse)
```

## Real data 

From https://github.com/chvlyl/PLEASE

Inflammation, Antibiotics, and Diet as Environmental Stressors of the Gut Microbiome in Pediatric Crohn’s Disease. Cell Host & Microbe. 2015


### Load the raw data

The "unclassfied" taxa were removed and the total relative abundance in each sample were normalized to be one. "P","F","G","S" at the beginning of each file indicate taxonomic levels "phylum", "family", "genus", "species".

We use genus abundance data.
```{r}
PLEASE.raw = 
  read_excel("./real data/G_Remove_unclassfied_Renormalized_Merge_Rel_MetaPhlAn_Result.xlsx",
             col_names = T) %>% 
  as.data.frame() 
row.names(PLEASE.raw) = PLEASE.raw[,1]
PLEASE.raw = PLEASE.raw[,-1]
colname = as.numeric(colnames(PLEASE.raw))
samplepoint = as.Date(colname,origin = "1900-01-01")
head(PLEASE.raw[,1:5])
```


```{r}
taxa.raw <- data.frame(t(PLEASE.raw),
                       row.names = strtrim(samplepoint,7))

### Make sure you load the data correctly
cat('samples','taxa',dim(taxa.raw),'\n')
taxa.raw[1:3,1:3]
```

Each row represent a sample.


### Load total non-human read counts
```{r}
human.read.file <- './real data/please_combo_human_reads.xlsx'
human.read <- 
  read_excel(human.read.file, col_names = T) %>% 
  as.data.frame() %>% 
  mutate(
    Sample = strtrim(as.Date(Sample,origin = "1900-01-01"),7)
  )

head(human.read)

## first column: sample id
```

```{r}
### Filter low depth samples (low non human reads)
low.depth.samples <- subset(human.read,NonHumanReads<10000)
head(low.depth.samples[,1:5])
```


```{r}
### Delete these samples from PLEASE data.
# row.names(taxa.raw)
# row.names(low.depth.samples)
row.names(taxa.raw)[which(row.names(taxa.raw) %in% low.depth.samples$Sample)]
### Before deletion
dim(taxa.raw)
### After deletion
taxa.raw <- taxa.raw[-which(rownames(taxa.raw) %in% low.depth.samples$Sample),]
dim(taxa.raw)
```

#### Filter low abundant bacterial data
```{r}
### Filter low abundant bacterial data
filter.index1 <- apply(taxa.raw,2,function(X){sum(X>0)>0.4*length(X)})
filter.index2 <- apply(taxa.raw,2,function(X){quantile(X,0.9)>1})
taxa.filter <- taxa.raw[,filter.index1 & filter.index2]
taxa.filter <- 100*sweep(taxa.filter, 1, rowSums(taxa.filter), FUN="/")
cat('after filter:','samples','taxa',dim(taxa.filter),'\n')
cat(colnames(taxa.filter),'\n')
head(rowSums(taxa.filter))
```

After filter, there remains 18 bacteria in the taxa table. 

```{r}
### 
taxa.data <- taxa.filter
dim(taxa.data)
```


### Load sample information
```{r include=FALSE}
# The processed clinical data and sample information

# sample.info.file <- 'https://raw.githubusercontent.com/chvlyl/PLEASE/master/1_Data/Processed_Data/Sample_Information/2015_02_13_Processed_Sample_Information.csv'
# sample.info <- read.csv(sample.info.file,row.names=1)

sample.info.file = "real data/2015_02_13_Processed_Sample_Information.xlsx"
sample.info <- 
  read_excel(sample.info.file,col_names = T) %>% 
  as.data.frame()
colnames(sample.info) = replace(colnames(sample.info),1,"Sample")
```

```{r}
head(sample.info)
```


#### create covariates, Time, Treatment(antiTNF+EEN)
```{r}
complete_subject =
  sample.info %>% 
  filter(Sample %in% rownames(taxa.data)) %>%
  filter(Treatment.Specific!='PEN')%>%
  dplyr::select(Sample,Time,Subject,Response,Treatment.Specific) %>%
  group_by(Subject) %>% 
  summarise(count = n()) %>% 
  filter(count==4) 

reg.cov = 
  sample.info %>% 
  filter(Subject %in%complete_subject$Subject) %>% 
  mutate(Treat=ifelse(Treatment.Specific=='antiTNF',1,0)) %>%
  dplyr::mutate(Subject=paste('S',Subject,sep='')) %>%
  dplyr::mutate(Time=ifelse(Time=='1',0,ifelse(Time=='2',1,ifelse(Time=='3',4,ifelse(Time=='4',8,NA))))) %>%
  dplyr::mutate(Time.X.Treatment=Time*Treat) %>%
  dplyr::select(Sample,Subject,Time,Response,Treat,Time.X.Treatment,everything())
```


### take out first time point

```{r}
reg.cov.t1   <-  subset(reg.cov,Time==0)
rownames(reg.cov.t1) <- reg.cov.t1$Subject

reg.cov.t234 <-  subset(reg.cov,Time!=0)
reg.cov.t234 <- data.frame(
  baseline.sample=reg.cov.t1[reg.cov.t234$Subject,'Sample'],
  baseline.subject=reg.cov.t1[reg.cov.t234$Subject,'Subject'],
  reg.cov.t234,
  stringsAsFactors = FALSE)

head(reg.cov.t234)
```




```{r}
taxa_all = colnames(taxa.data)
store = function(taxa){
# X: Baseline abundance time Treat
# Y: Response abundance at time 1 4 8
X <- data.frame(
      Baseline=taxa.data[reg.cov.t234$baseline.sample,taxa]/100,
      reg.cov.t234[,c('Time','Treat')])

rownames(X) <- reg.cov.t234$Sample
Y <- taxa.data[reg.cov.t234$Sample, taxa]/100

return(list(X = X,  
            Y = Y))
}


store_results = lapply(taxa_all, store)

#example
store_results[[1]]$X
```

